Learning and comparing functional connectomes across subjects

### 2013-01-03 paper.tex
iter

For the problem of recovering the functional-connectivity
\emph{structure}, \emph{i.e.}\ finding which region is connected to
which, sparse inverse covariance estimators have been found to be
% CAMERON Isn't there a paper from Stanford about this as well??
efficient \cite{varoquaux2010c, smith2011}. The intuition for relying on
inverse covariance rather than correlation stems from that fact that
standard correlation (marginal correlation) between two variables $a$ and
$b$ also capture the effects of other variables: strong correlation of
$a$ and $b$ with a third variable $c$ will induce a correlation between
$a$ and $b$. On the opposite, the inverse covariance\footnote{Covariance
and correlation matrices differ simply by the fact that a covariance
matrix captures the amplitude of a signal, via its variance, while a
correlation matrix is computed on standardized (zero mean, unit variance) signals.}
matrix captures
partial correlations, removing the effect of other variables
\cite{marrelec2006a}. In the small sample limit, this removal is
challenging from the statistical standpoint, and an assumption of sparsity,
\emph{i.e.}\ that only few variables need to be considered at a time, is
important to estimate a good inverse covariance. Various estimation
strategies exist for sparse inverse covariance, and have an impact on the
resulting networks \cite{varoquaux2012,varoquaux2010c}. The
Graph Lasso ($\ell_1$-penalized maximum-likelihood estimator)
\cite{friedman2008} is in general a good approach for structure recovery. In group studies,
the $\ell_{21}$ estimator \cite{varoquaux2010c,honorio2012} is useful to
impose a common sparsity structure across different subjects and achieve
better recovery of this common structure. Simply put, these approaches
are necessary because estimation noise creates a background structure
(see fig.\,\ref{fig:icov_estimators}); however, unlike in a univariate situation, the
parameters are not independent, and the spurious background connections
degrade the estimation of the actual connections. The sparse estimators
make a compromise between imposing simpler models, \emph{i.e.}\ with less
% CAMERON "fitting well the data" needs to be fixed
connections, and fitting well the data. This compromise is set via a
regularization parameter which controls the sparsity of the estimate. A
good rule of thumb to choose this parameter is via cross-validation
\cite{varoquaux2010c}.

@@

% A bit of technical details
% technical details
Technically, predictive modeling is a supervised machine learning problem
where a target to be predicted --\emph{e.g.}\ age, disease state,
cognitive state-- is available for each observation of the data. In the
context of comparing connectomes, features used in the predictive model
usually correspond to bivariate measures of connectivity, or any other
summary metrics discussed earlier.  
%
The quality of a predictive model is determined by its prediction accuracy (or
generalization ability) which is measured using one or more iterations of
cross-validation. Cross-validation begins by seperating the available data into
independent training and testing subsets.  A predictive model is learned from
the training data and applied to the testing data to predict corresponding
values of the target variable. Prediction accuracy is then assessed by comparing
these predictions to the observed value of the target variable. This procedure
can be iterated several times for different splits of the data and averaged to
form a single measure of prediction accuracy.  One of the most important
parameters for cross validation is the manner in which data are subdivided.
There are several strategies for constructing the training dataset ranging from
using all but one observations (leave-one-out cross-validation), using a
fraction of the observations (k-fold cross validation; kfcv), to resampling the
observations with replacement (boostrap cross-validation, bscv). Although loocv
has been popular in neuroimaging applications of MVPA since it makes the largest
amount of data available for training, its estimates are more variable than kfcv
or bscv [tibs, book]. Split-half cross validation (two-fold cross validation,
2fcv), minimizes the estimation variance, at the cost of using less training
data, and also permits the simultaneous estimation of model reproducibility
using the NPAIRs framework []. Bootstrap cross-validation minimizes estimation
variance while maximising the amount of data available for training at the cost
of introducing bias. This bias is due to the fact that on average 63.2\percent
of the testing data overlaps with the training data, the .632\plus bootstrap
corrects for this bias.  No matter which cross validation scheme is used, the
accurate estimation of generalization ability requires that the testing data and
training data are independent and identically distributied (iid). The meaning of
independence here depends on the context in which the predictive model will be
applied. For example, if observations correspond to subjects, then testing with
scans acquired from the same subject, even acquired at different points in time,
will introduce bias. If observations are different time points for the same
subject, with the goal of testing whether the model can predict a later time
point for that subject, then scans acquired from a later time point for that
subject may be used for testing as long as enough time has passed to account for
temporal autocorrelation.

>>>>

For the problem of recovering the functional-connectivity
\emph{structure}, \emph{i.e.}\ finding which region is connected to
which, sparse inverse covariance estimators have been found to be
% CAMERON Isn't there a paper from Stanford about this as well??
% GAEL: I cannot find it with a quick Google. Do you have a name?
efficient \cite{varoquaux2010c, smith2011}. The intuition for relying on
inverse covariance rather than correlation stems from that fact that
standard correlation (marginal correlation) between two variables $a$ and
$b$ also capture the effects of other variables: strong correlation of
$a$ and $b$ with a third variable $c$ will induce a correlation between
$a$ and $b$. On the opposite, the inverse covariance\footnote{Covariance
and correlation matrices differ simply by the fact that a covariance
matrix captures the amplitude of a signal, via its variance, while a
correlation matrix is computed on standardized (zero mean, unit variance) signals.}
matrix captures
partial correlations, removing the effect of other variables
\cite{marrelec2006a}. In the small sample limit, this removal is
challenging from the statistical standpoint, and an assumption of sparsity,
\emph{i.e.}\ that only few variables need to be considered at a time, is
important to estimate a good inverse covariance. Various estimation
strategies exist for sparse inverse covariance, and have an impact on the
resulting networks \cite{varoquaux2012,varoquaux2010c}. The
Graph Lasso ($\ell_1$-penalized maximum-likelihood estimator)
\cite{friedman2008} is in general a good approach for structure recovery. In group studies,
the $\ell_{21}$ estimator \cite{varoquaux2010c,honorio2012} is useful to
impose a common sparsity structure across different subjects and achieve
better recovery of this common structure. Simply put, these approaches
are necessary because estimation noise creates a background structure
(see fig.\,\ref{fig:icov_estimators}); however, unlike in a univariate situation, the
parameters are not independent, and the spurious background connections
degrade the estimation of the actual connections. The sparse estimators
make a compromise between imposing simpler models, \emph{i.e.}\ with less
% CAMERON "fitting well the data" needs to be fixed
connections, and fitting well the data. This compromise is set via a
regularization parameter which controls the sparsity of the estimate. A
good rule of thumb to choose this parameter is via cross-validation
\cite{varoquaux2010c}.

@@

% A bit of technical details
% technical details
Technically, predictive modeling is a supervised machine learning problem
where a target to be predicted --\emph{e.g.}\ age, disease state,
cognitive state-- is available for each observation of the data. In the
context of comparing connectomes, features used in the predictive model
usually correspond to bivariate measures of connectivity, or any other
summary metrics discussed earlier.  
%
The quality of a predictive model is determined by its prediction accuracy (or
generalization ability) which is measured using one or more iterations of
cross-validation. Cross-validation begins by seperating the available data into
independent training and testing subsets.  A predictive model is learned from
the training data and applied to the testing data to predict corresponding
values of the target variable. Prediction accuracy is then assessed by comparing
these predictions to the observed value of the target variable. This procedure
can be iterated several times for different splits of the data and averaged to
form a single measure of prediction accuracy.  One of the most important
parameters for cross validation is the manner in which data are subdivided.
There are several strategies for constructing the training dataset ranging from
using all but one observations (leave-one-out cross-validation), using a
fraction of the observations (k-fold cross validation; kfcv), to resampling the
observations with replacement (boostrap cross-validation, bscv). Although loocv
has been popular in neuroimaging applications of MVPA since it makes the largest
amount of data available for training, its estimates are more variable than kfcv
or bscv [tibs, book]. Split-half cross validation (two-fold cross validation,
2fcv), minimizes the estimation variance, at the cost of using less training
data, and also permits the simultaneous estimation of model reproducibility
using the NPAIRs framework []. Bootstrap cross-validation minimizes estimation
variance while maximising the amount of data available for training at the cost
of introducing bias. This bias is due to the fact that on average 63.2\%
of the testing data overlaps with the training data, the .632+ bootstrap
corrects for this bias.  No matter which cross validation scheme is used, the
accurate estimation of generalization ability requires that the testing data and
training data are independent and identically distributied (iid). The meaning of
independence here depends on the context in which the predictive model will be
applied. For example, if observations correspond to subjects, then testing with
scans acquired from the same subject, even acquired at different points in time,
will introduce bias. If observations are different time points for the same
subject, with the goal of testing whether the model can predict a later time
point for that subject, then scans acquired from a later time point for that
subject may be used for testing as long as enough time has passed to account for
temporal autocorrelation.
